### 2020-05-14 manuscript.Rmd
changes to introduction

Insufficient physical activity is a leading cause of disease and disability worldwide.[@who2009] For the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic background are disproportionately affected by physical inactivity-related diseases. Increasing the physical activity levels of the population is therefore high on the public health agenda: it not only has the potential to improve quality of life, reduce mortality rates and alleviate the strain on health and social care services, but also reduce the gap in health inequalities.

@@

parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professional with respect to both. Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation particularly in individuals who were previously inactive.[@stevinson2004; @wiltshire2018]

Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely driven by *demand*, not *need*. It might therefore be the case that parkrun events are primarily located in areas that are relatively well-off anyway, while people living in more deprived communities may not have the same opportunities to participate. In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@sportengland2018]

The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England is equitable across areas with different levels of deprivation. Secondly, to identify 200 optimal locations for future events, in order to improve geographic access for deprived communities.

>>>>

Insufficient physical activity is a leading cause of disease and disability worldwide.[@who2009] In the UK, around one in six deaths is attributable to low levels of physical activity.[@lee2012] It is also a major contributor to health inequalities, as people from low socio-economic background are more likely to be inactive [@farrell2014socioeconomic] and are disproportionately affected by physical inactivity-related diseases [@tanaka2018socioeconomic]. Increasing the physical activity levels of the population is therefore high on the public health agenda: it not only has the potential to improve quality of life, reduce mortality rates and alleviate the strain on health and social care services, but also reduce the gap in health inequalities.

@@

parkrun, an international movement which organises free weekly 5km running events, might provide valuable lessons for public health professional with respect to both. Since its founding in 2004, as a small event in London with 13 participants, it has grown to become one of the world's largest mass participation sporting events, with up to 360,000 participants in more than 20 countries.[@ref] The volunteer-led running events are often characterised as accessible and inclusive.[@stevinson2015] The organisation has been widely praised as being successful in encouraging participation, particularly in individuals who were previously inactive or did not meet public health guidelines.[@stevinson2004; @wiltshire2018]

Notwithstanding these subjective accounts, the expansion of parkrun in England, as elsewhere, has been largely grassroots, driven by *demand* rather than *need*. It might therefore be the case that parkrun events are primarily located in areas that are relatively affluent, while people living in more deprived communities may not have the same opportunities to participate. In 2019, Sport England announced funding to support the creation of 200 new parkrun events across England within three years, with the specific aim of increasing participation of individuals from lower socio-economic groups.[@sportengland2018]

The aims of this study are two-fold: firstly, to evaluate whether geographic access to parkrun events in England is equitable across areas with different levels of deprivation. Secondly, to identify 200 optimal locations for future events in order to improve geographic access, in particular for deprived communities.
